Score-based tests of measurement invariance: Use in practice

### 2014-02-10 frontiers/frontiers.Rnw
small modifications and proposed change in response 2.6
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@313 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
Minor changes to ms in response to comment 2.6.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@314 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

The family of tests described here have recently been applied
to the study of measurement invariance \cite{MerZei13,MerFanZei}, though their
practical application has been limited to a small set of simulations and
data examples.  In this paper, we provide a detailed illustration of
the tests' use and performance under
scenarios likely to be encountered in practice.  While the previous
papers have described and studied the tests under ideal conditions, we
focus here on two topics of interest to the applied researcher:
software considerations for carrying out the tests, and test
performance under model misspecification.
%In order to fix the psychometric scale, we need to specify the violating parameter.  
%Moreover, in practice model misspecification is an inevitable issue to a certain extent 
%and may interfere with parameter invariance specificity. A family of tests derived from 
%stochastic process has been recently proposed to identify specific parameters impacted 
%by measurement invariance violation. However, these tests have only been used to a limited 
%type of parameter violation. It is this paper's intent to provide a detailed examination 
%of the tests' performance under practical research scenarios. 
In the following sections, we first briefly review the  theoretical framework of
the proposed tests and provide a short tutorial illustrating the use of the
tests in \proglang{R} \cite{R11}. Next, we study the tests' performance in simulations that mimic
practical research scenarios.  Finally, we provide some further discussion on
the tests' use in practice.  We acknowledge 
problems associated with $p$-values (e.g., corrections for multiple
tests, researcher degrees of freedom, interpretations, etc) but
ignore them throughout the paper for expository purposes.

@@

In this section, we demonstrate how the above tests can be carried out in \proglang{R}, 
using the package \pkg{lavaan} \cite{lavaan11} for model estimation and \pkg{strucchange} 
\cite{ZeiLei02,Zei06} for testing.  We use data from \citeA{FroFan11} concerning 
the applicability of adult gratitude scales to youth, available in the~\proglang{R} 
package \pkg{psychotools} \cite{zeilei12}.  The data consist of responses to three adult 
gratitude scales from $n=1401$ youth aged 10--19 years. 
The original authors 
were specifically interested in whether the scales were measurement invariant 
across age.  Because the sample size at each age was unbalanced, the authors 
created age groups of approximately equal sample size.  In the examples below, 
we test for measurement invariance across these age groups.

>>>>

The family of tests described here have recently been applied
to the study of measurement invariance \cite{MerZei13,MerFanZei}, though their
practical application has been limited to a small set of simulations and
data examples.  In this paper, we provide a detailed illustration of
the tests' use and performance under
scenarios likely to be encountered in practice.  While the previous
papers have described and studied the tests under ideal conditions, we
focus here on two topics of interest to the applied researcher:
software considerations for carrying out the tests, and test
performance under model misspecification.
%In order to fix the psychometric scale, we need to specify the violating parameter.  
%Moreover, in practice model misspecification is an inevitable issue to a certain extent 
%and may interfere with parameter invariance specificity. A family of tests derived from 
%stochastic process has been recently proposed to identify specific parameters impacted 
%by measurement invariance violation. However, these tests have only been used to a limited 
%type of parameter violation. It is this paper's intent to provide a detailed examination 
%of the tests' performance under practical research scenarios. 
In the following sections, we first briefly review the  theoretical framework of
the proposed tests and provide a short tutorial illustrating the use of the
tests in \proglang{R} \cite{R11}. Next, we study the tests' performance in simulations that mimic
practical research scenarios.  Finally, we provide some further discussion on
the tests' use in practice.
% We acknowledge problems associated with $p$-values (e.g., corrections for multiple
% tests, researcher degrees of freedom, interpretations, etc) but
% ignore them throughout the paper for expository purposes.

@@

In this section, we demonstrate how the above tests can be carried out in \proglang{R}, 
using the package \pkg{lavaan} \cite{lavaan11} for model estimation and \pkg{strucchange} 
\cite{ZeiLei02,Zei06} for testing.  We use data from \citeA{FroFan11} concerning 
the applicability of adult gratitude scales to youth, available in the~\proglang{R} 
package \pkg{psychotools} \cite{zeilei12}.  The data consist of responses to three adult 
gratitude scales from $n=1401$ youth aged 10--19 years. 
The original authors 
were specifically interested in whether the scales were measurement invariant 
across age.  Because the sample size at each age was unbalanced, the authors 
created age groups of approximately equal sample size.  In the examples below, 
we test for measurement invariance across these age groups.
For illustrative purposes, we conduct multiple tests and
compare them to the traditional significance level of .05.  In
practice, however, one should generally adjust the significance level
for the number of tests carried out.
